Mapping factors associated with adolescents' emotion dysregulation: A text-mining based systematic review

### 2020-10-18 manuscript.Rmd
Submit to DPP

---
title             : "Mapping factors associated with adolescents' emotion dysregulation: A text-mining based systematic review"
shorttitle        : "ADOLESCENT EMOTION DYSREGULATION"

@@

authornote: |
  This work is supported by a NWO Veni grant (NWO
  grant number VI.Veni.191G.090). Acknowledgement: Lukas Beinhauer contributed
  substantially to the screening of articles.

abstract: |
  Test me

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["veni_sysrev.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
knit              : worcs:::cite_essential
---

```{r setup, include = FALSE}
library("papaja")
library(bibliometrix)
library(revtools)
library(bib2df)
```
<!-- Reviewer suggestions:
Kristin Buss -->

@@

<!-- Emotion regulation in adolescence is important -->
A key developmental challenge in adolescence is acquiring mature emotion
regulation skills [@croneUnderstandingAdolescencePeriod2012;
@zimmermannEmotionRegulationEarly2014]. Emotion regulation is integral to
mental health[@aldaoEmotionregulationStrategiesPsychopathology2010;
@schaferEmotionRegulationStrategies2017; @braetEmotionRegulationChildren2014]
and social functioning [@reindlSocializationEmotionRegulation2016]. For as many
as one in five people, adolescence marks the onset of emotion
regulation-related mental illness, which can persist throughout the life course
[see @leeAdolescentMentalHealth2014]. It is therefore crucial to identify which
risk factors and environmental hazards render some adolescents more susceptible
to emotional difficulties than others. Although there is an abundance of
empirical work on adolescents' emotion regulation, the literature is somewhat
fragmented, as the topic has been investigated from different (sub)disciplines.
This study seeks to overcome this limitation by mapping all factors currently
thought to be relevant to adolescent emotion regulation, based on a systematic
review and bibliographic analysis. <!-- Ergens anders: Similarly, although
there are several *relevant* theories, there is no specific theoretical
framework to guide research on adolescents' emotion regulation. Although theory
development is beyond the scope of this paper, we set out to generate a
proto-theoretical nomological network of the potential risk factors,
manifestations, and outcomes of emotion regulation in adolescence. We discuss
how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.-->

@@

Adolescence is defined as a life phase ranging from pubertal onset to
adult-like independence [@steinbergAgeOpportunityLessons2014]. This definition
invokes both biological- and socio-cultural factors. It has been argued that,
due to accellerated pubertal onset and delayed transitions to adult roles,
adolescence now ranges from 10-24 years in the modern Western world
[@sawyerAgeAdolescence2018].

@@

Adolescence is increasingly seen as a developmentally sensitive period for
mature emotion regulation skills. During this period, children experience rapid
biological, cognitive, and social changes that prompt new emotional
experiences, and tax existing regulatory abilities
[@steinbergAgeOpportunityLessons2014]. At the same time, the staggered
development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap, where youngsters pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes
[@croneUnderstandingAdolescencePeriod2012]. Due to these coalescing changes,
adolescents display a restricted repertoire of emotion regulation strategies
compared to younger or older individuals
[@zimmermannEmotionRegulationEarly2014]. Consequently, adolescents experience
more frequent and intense (negative) emotions than children or adults [see
@silkAdolescentsEmotionRegulation2003]. The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially. These changes
render adolescents temporarily vulnerable to emotional dysregulation, but
ultimately serve a functional role in acquiring mature emotion regulation
skills. Eventually, most adolescents emerge ready to take on the challenges of
early adulthood. Unfortunately, a sizeable proportion of adolescents instead
manifests enduring emotional difficulties.

@@

<!--Acquiring mature emotion regulation abilities is important, because t-->The
consequences of difficulties in emotion regulation are extensive,
well-documented, and persist throughout the life course. Meta-analyses indicate
that such difficulties are associated with externalizing, and more strongly,
with internalizing psychopathology
[@aldaoEmotionregulationStrategiesPsychopathology2010;
@schaferEmotionRegulationStrategies2017]. Individuals who engage in maladaptive
emotion regulation strategies also experience more negative emotions,
diminished well-being, and greater strain in interpersonal relationships
[@grossIndividualDifferencesTwo2003a; @bellRelationshipsInputsOutputs2000].
<!--Adolescence also appears to be a key period for developing mature emotion
regulation skills - but not all adolescents do so successfully. As many as one
in five adolescents instead develop severe emotional difficulties.--> Given the
prevalence of emotional difficulties and their extensive implications for
individual mental health, wellbeing, and social functioning, and downstream
cost to society as a whole [@leeAdolescentMentalHealth2014], it is essential to
have a solid understanding of the risk factors associated with adolescents'
emotion regulation difficulties.

@@

Theory is of paramount importance when seeking to understand or explain a
phenomenon. Although there are many relevant theories, none have specifically
addressed emotion regulation development in adolescence. Several recent
publications have undertaken the monumental task of comprehensively reviewing
theory relevant to emotional development
[@bussTheoriesEmotionalDevelopment2019;
@riedigerEmotionRegulationAdolescence2014]. Thanks, in part, to their efforts,
we can provide a summary review of the theoretical landscape as context for our
empirical contribution.<!--Several recent publications have undertaken the
monumental task of providing a comprehensive overview of theories of emotional
development [e.g., @bussTheoriesEmotionalDevelopment2019;
@holodynskiDevelopmentEmotionsEmotion2006;
@riedigerEmotionRegulationAdolescence2014], and of the empirical work
pertaining to emotions in adolescence
[@coe-odessEmergentEmotionsAdolescence2019].  -->
Two theories widely invoked to frame research on development, including
emotional development, are Bronfenbrenner's bioecological model, and Sameroff's
transactional model. Bronfenbrenner's model
[@bronfenbrennerBioecologicalModelHuman2007] describes how the environment
shapes individual development. Each individual is imbued with certain
biological predispositions, and develops over time, in interaction with
contextual influences. These influences range from the microsystem, composed of
those people close to the individual, to the macrosystem, consisting of
indirect political and economic influences, to the exosystem, consisting of
cultural norms and values. Sameroff's model is somewhat compatible with the
bioecological model [@sameroffUnifiedTheoryDevelopment2010], but focuses more
on the interplay between nature and nurture. It conceptualizes development as a
product of reciprocal influences between the child and the environment.
Sameroff distinguishes between proximal influences, for instance, through
parents (the microsystem in Bronfenbrenner's model), and distal influences;
structural factors indirectly shaping development, like socio-economic status,
schools, and the community (macro- and exosystem). With increasing age, distal
influences gradually gain ground on proximal influences.

Turning now towards theories specifically addressing emotion regulation in
adolescence, Hall's notion of "storm and stress" is one of the oldest [see
@arnettAdolescentStormStress1999]. It describes how hormonal changes diminish
self-control and increase reactivity, leading to difficulties in emotion
regulation, conflict with parents, and risky behavior. Recent work provides a
more nuanced perspective on this theory. For example, although the notion of
diminished self-control and increased emotional reactivity has been upheld, it
has been recast as a necessary change that facilitates emotional maturation
[@croneUnderstandingAdolescencePeriod2012], at the risk of emotional
disturbance [@arnettAdolescentStormStress1999; @leeAdolescentMentalHealth2014].
More recently, Sroufe [-@sroufeEmotionalDevelopmentOrganization1995] presented
a theory focused on normative emotional development in early childhood. It
argues that, as children grow older, their increasing self-regulatory abilities
drive a transition from external emotion regulation by primary caregivers
towards autonomous emotion regulation. This theory focuses on two drivers of
development: social and cognitive influences. Social influences mainly occur
through parents' co-regulation, parenting behaviors, and parent-child
attachment. Cognitive influences occur through the development of the central
nervous system (CNS), cognition, and self-regulation. Several other theories,
introduced below, have extemporized on these sources of influence.

The emphasis on neorocognitive development as a driver of emotional
development, including emotion regulation development, is echoed in several
other models. For example, polyvagal theory closely links emotional experience
- and regulation - to autonomous nervous system functioning
[@porgesOrientingDefensiveWorld1995]. A more recent neorocognitive theory, with
substantial relevance to adolescents' emotion regulation development, is Crone
and Dahl's model of social-affective engagement and goal flexibility
[@croneUnderstandingAdolescencePeriod2012]. This theory is based on the notion
of a "maturity gap" in middle-adolescence, arising from a developmental
asymmetry between motivational and inhibitory brain circuits
[@craccoEmotionRegulationChildhood2017;
@caseyBrakingAcceleratingAdolescent2011]. However, it sets out to explain - to
a greater extent than other writings - adolescents' diverging destinies; the
question why some youngsters flourish while others languish. Crone and Dahl
argue that adolescents' cognitive engagement is dynamically responsive to
social and motivational goal salience. This flexibility, on the one hand,
prepares adolescents to effectively engage cognitive systems in novel
challenging situations in a way that facilitates developing mature regulatory
abilities. On the other hand, it places them at risk to act impulsively in
pursuit of peer approval. This theory focuses primarily on cognitive factors
and the role of peers, with less attention to factors such as parenting.

@@

Aside from theories of emotional development, theories of the phenomenon of
emotion regulation are also somewhat relevant. One influential theory is Gross'
[-@grossHandbookEmotionRegulation2013] process model, which describes the
process of emotion regulation, from eliciting cue to ultimate response.
Individuals modulate the different stages of this process using strategies,
consciously or otherwise. The effectiveness, desireability, and consequences of
different strategies depend on the cultural context [see
@bariolaChildAdolescentEmotion2011]. Similar to Gross' theory, the social
information processing theory of emotion also describe the role of intrapsychic
processes and strategies in emotion experience and reulation
[@lemeriseIntegratedModelEmotion2000].

@@

## Shortcomings of the theory

Existing theoretical models have several limitations. One crucial shortcoming
is the lack of explicit attention to the life phase of adolescence. This is
unfortunate, as empirical research consistently indicates that adolescence is a
crucial life phase for emotion regulation development. Adolescence differs
qualitatively from infancy, childhood, and adulthood. It is therefore
insufficient to simply extrapolate theories focused on these ages. Other
theories do focus on adolescence specifically [e.g.,
@croneUnderstandingAdolescencePeriod2012], but only address emotion regulation
in passing. A second, related, shortcoming is that some relevant theories lack
a developmental component <!--, and thus lack a well-substantiated
understanding of developmental processes--> [see
@bussTheoriesEmotionalDevelopment2019;
@croneUnderstandingAdolescencePeriod2012]. A third shortcoming is that theories
vary widely in scope: Some are very broad and all-encompassing, but somewhat
non-specific, such as the biopsychosocial models of Bronfenbrenner and
Sameroff. Others describe a specific phenomenon in great detail, but are not
embedded in a larger unifying framework. Examples include Morris' tripartite
model, Crone and Dahl's model, and polyvagal theory. It would be beneficial to
bridge these two levels of theory. One limitation echoed across many
publications in this field is that more theory formation is required to provide
a unified framework that could guide future empirical work. The present study
seeks to aid such theory development, by providing an inductive conceptual
review of the field.

@@

While theory provides a top-down understanding of an area of research,
literature reviews can provide a bottom-up, inductive understanding. Given the
noted absence of a single overarching theory, reviews might provide additional
insight into factors considered relevant in the etiology of adolescents'
emotion regulation competencies. For example, a review by
@bariolaChildAdolescentEmotion2011 addresses individual psychological factors
such as temperament, and biological factors like neurocognitive development and
genetic predispositions. Furthermore, they address proximal social influences,
including parents, teachers, and peers, and specific mechanisms through which
these individuals exert their influence, including parenting and modeling.
Finally, they address distal factors, including the role of culture and the
media. Barriola and colleagues also offer recommendations for future research,
<!--most notably, they call for a robust theoretical framework to guide future
research.--> calling for research on the role of parents beyond early
childhood, and including the role of fathers - a limitation which has since
been addressed by a new wave of empirical research on fathers' role in
emotional development [@vanlissaRoleFathersMothers2019;
@vanlissaMothersFathersQuantitative2020].

@@

There are notable parallels between the relevant factors highlighted by these
narrative reviews, and the aforementioned theoretical literature. Nevertheless,
these literature reviews also touch upon issues that have received little
attention in theoretical work. This highlights the value of both top-down and
bottom-up approaches to understanding the phenomenon of adolescents' emotion
regulation. One remaining limitation is that, to date, all reviews in this
field have been unstructured narrative reviews. The present publication seeks
to further complement preceding work by conducting a comprehensive empirical
literature review, and mapping the factors associated with adolescent emotion
regulation.

@@

The present paper aims to develop a nomological network of the constructs
related to adolescent emotion dysregulation, based on a systematic review of
the literature. A nomological network is a diagrammatic representation of a
phenomenon; a proto-theoretical device that describes relationships between
constructs relevant to the theory [see @alaviAligningTheoryMethodology2018]. It is a precursor to a more
explicit theory of the phenomenon. The present paper uses an inductive approach
to develop this nomological network. First, we conduct a systematic search to
elicit a corpus of relevant literature. Second, we use text mining techniques
to extract relevant constructs from the corpus. Third, we use a dictionary to
pare down the extracted constructs to meaningful superordinate categories.
Fourth, we map interrelations between these constructs. Fifth, we classify
contstructs as potential predictors or outcomes, and we categorize predictors
by level of analysis<!-- (CJ: Sameroff, which level of biopsychosocial)-->.
<!-- t, using text mining to map potential predictors of adolescent emotion
dysregulation.
identify relevant constructs of adolescents’ emotion regulation development,
and moderators of parents’ role therein. The resulting insights are
consolidated into theory.
The present study intends to map the extant literature
First, I will systematically review predictors currently considered relevant in
the adolescent emotion regulation literature.  -->

@@

if(!file.exists("recs_final.csv")){
  recs_final <- unique_recs[screen$key[!screen$drop_these], ]
  write.csv(recs_final, "recs_final.csv", row.names = FALSE)
} else {
  recs_final <- read.csv("recs_final.csv")
}
```
Papers were screened based on two main criteria:
First, papers had to address emotion regulation, emotional problems, or similar.
Second, the target population must overlap (at least in part) with adolescence, defined as the age range from 10-25.
Thus, for example, an article reviewing the role of hormonal changes in emotional adjustment throughout the entire life course, including adolescence, would be deemed relevant.
Screening was initially conducted in Rayyan QRCI [@Ouzzani2016].
Papers were sorted using Rayyan's ranking algorithm.
To train the algorithm on exclusion criteria, screening focused on abstracts most likely to be excluded.
After coding `r sum(screen$rayyan == "True")+sum(screen$rayyan == "False")` papers, screening continued in ASReview - a free, open source alternative to Rayyan QRCI [@van_de_schoot_rens_2020_3828293].
This program uses text mining and active learning to build a customizable machine learning model for article inclusion.
We used a naive Bayes model, and focused on exclusion of irrelevant papers.
In ASReview, an additional `r sum(table(screen$included, screen$rayyan)[, 1])` papers were screened.
Screening again focused on articles most likely to be excluded, and continued until, among the most recently screened 100 papers, only 6 were excluded.
The screening procedure is detailed in Figure \@ref(prismachart).
In total, `r nrow(recs_final)` papers were retained for analysis.

@@

## Corpus description

```{r, eval = FALSE}
# Plot papers per year
df_plot <- as.data.frame(table(recs_final$PY), stringsAsFactors = FALSE)
df_plot$Year = as.numeric(df_plot$Var1)
ggplot(df_plot, aes(x= Year, y= Freq))+geom_point() + geom_path(group =1)+
  theme_bw() + scale_y_log10()
```

# Analysis 1: Keyword mapping

The corpus for this first analysis consisted of the author-provided keywords for the selected articles.
We extracted keywords by document, and used a dictionary approach
to classify related terms.
The dictionary was created specifically for this study.
The lead author composed a first draft of the dictionary based on prior knowledge of the field.
This draft was subsequently updated in two ways:
First, while writing the introduction for this paper, key terms from the theoretical and empirical literature were added to the dictionary.
Second, by including commonly occurring uncategorized words during preliminary analyses of the data.
An exclusion list of methodological terms and other non-substantive words was generated using the same procedure.

>>>>

---
title             : "Mapping correlates of adolescent emotion dysregulation: A text-mining based systematic review"
shorttitle        : "ADOLESCENT EMOTION DYSREGULATION"

@@

authornote: |
  This work is supported by a NWO Veni grant (NWO
  grant number VI.Veni.191G.090). Acknowledgement: Lukas Beinhauer contributed
  substantially to the screening of articles. Conflicts of Interest: None.

abstract: |
  For many, adolescence marks the onset of emotion dysregulation and related mental health problems. It is therefore essential to have a solid grasp of relevant risk factors. Despite an abundance of empirical research, the literature is somewhat fragmented, as the topic has been investigated from different perspectives across subdisciplines. This systematic review uses text mining to map factors related to adolescent emotion dysregulation. A systematic search for publications that addressed emotion regulation in participants between the ages of 10-25 yielded 6305 suitable papers. One analysis focused on author-provided keywords, and a second extracted sequences of adjectives and nouns from the abstracts. Terms were classified with a dictionary describing 105 terms using 434 regular expressions. Terms were coded as putative “Causes”, “Outcomes”, “Protective factors”, or “Indicators” of emotion dysregulation. We reported term/document frequency, and force-directed graphs of term co-occurrence. These results reﬂected constructs commonly featured in theory and narrative reviews, but also identified underrepresented terms. This illustrates how text mining review may complement narrative reviews. Importantly, most constructs were directly related to emotion dysregulation, with few interconnections between constructs – reflecting the fragmented literature. This research constitutes a first step towards integrating the literature, and may inspire future empirical and theoretical work.

keywords          : "emotion regulation, adolescence, systematic review, text mining, machine learning"
wordcount         : "5400"
bibliography      : ["veni_sysrev.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man,noextraspace"
output            : papaja::apa6_pdf
knit              : worcs:::cite_essential
---

```{r setup, include = FALSE}
library("papaja")
library(bibliometrix)
library(revtools)
library(bib2df)
library(knitr)
library(kableExtra)
library(cowplot)
run_everything = FALSE
```
<!-- Reviewer suggestions:
Kristin Buss -->

@@

<!-- Emotion regulation in adolescence is important -->
A key developmental challenge in adolescence is acquiring mature emotion
regulation skills [@croneUnderstandingAdolescencePeriod2012;
@zimmermannEmotionRegulationEarly2014]. Emotion regulation is integral to
mental health[@aldaoEmotionregulationStrategiesPsychopathology2010;
@schaferEmotionRegulationStrategies2017; @braetEmotionRegulationChildren2014]
and social functioning [@reindlSocializationEmotionRegulation2016]. For as many
as one in five people, adolescence marks the onset of emotion
regulation-related mental illness, which can persist throughout the life course
[see @leeAdolescentMentalHealth2014]. It is therefore crucial to identify which
risk factors and environmental hazards render some adolescents more susceptible
to emotional difficulties than others. Although there is an abundance of
empirical work on adolescents' emotion regulation, the literature is somewhat
fragmented, as the topic has been investigated from different (sub)disciplines.
This systematic review seeks to overcome this limitation by using
bibliographic analysis to map all factors currently
considered to be relevant to adolescent emotion regulation. <!-- Ergens anders: Similarly, although
there are several *relevant* theories, there is no specific theoretical
framework to guide research on adolescents' emotion regulation. Although theory
development is beyond the scope of this paper, we set out to generate a
proto-theoretical nomological network of the potential risk factors,
manifestations, and outcomes of emotion regulation in adolescence. We discuss
how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.-->

@@

Adolescence is defined as a life phase ranging from pubertal onset to
adult-like independence [@steinbergAgeOpportunityLessons2014].
It is thus demarcated by both biological- and socio-cultural factors.
It has been argued that,
due to accellerated pubertal onset and delayed transitions to adult roles,
adolescence now ranges from 10-24 years in the modern Western world
[@sawyerAgeAdolescence2018].

@@

Adolescence is increasingly seen as a developmentally sensitive period for
mature emotion regulation skills. During this period, children experience rapid
biological, cognitive, and social changes that prompt new emotional
experiences, and tax existing regulatory abilities
[@steinbergAgeOpportunityLessons2014]. At the same time, the staggered
development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap, where youngsters pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes
[@croneUnderstandingAdolescencePeriod2012].
Due to these coalescing changes,
adolescents display a restricted repertoire of emotion regulation strategies
compared to younger or older individuals
[@zimmermannEmotionRegulationEarly2014]. Consequently, adolescents experience
more frequent and intense (negative) emotions than children or adults [see
@silkAdolescentsEmotionRegulation2003]. The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially. These changes
render adolescents temporarily vulnerable to emotional dysregulation, but
ultimately serve a functional role in acquiring mature emotion regulation
skills. Eventually, most adolescents emerge ready to take on the challenges of
early adulthood. Unfortunately, a sizeable proportion of adolescents instead
manifests enduring emotional difficulties [@leeAdolescentMentalHealth2014].

@@

<!--Acquiring mature emotion regulation abilities is important, because t-->The
consequences of difficulties in emotion regulation are extensive,
well-documented, and persist throughout the life course. Meta-analyses indicate
that such difficulties are associated with internalizing, and to a lesser extent, externalizing, psychopathology
[@aldaoEmotionregulationStrategiesPsychopathology2010;
@schaferEmotionRegulationStrategies2017]. 
Individuals who engage in maladaptive
emotion regulation strategies also experience more negative emotions,
diminished well-being, and greater strain in interpersonal relationships
[@grossIndividualDifferencesTwo2003a; @bellRelationshipsInputsOutputs2000].
<!--Adolescence also appears to be a key period for developing mature emotion
regulation skills - but not all adolescents do so successfully. As many as one
in five adolescents instead develop severe emotional difficulties.--> 
Given the
prevalence of emotional difficulties and their implications for
individual mental health, wellbeing, and social functioning, and their downstream
cost to society as a whole [@leeAdolescentMentalHealth2014], it is essential to
have a solid understanding of the risk factors associated with adolescents'
emotion regulation difficulties.

@@

Theory is of paramount importance when seeking to understand or explain a
phenomenon. Although there are many relevant theories, none have specifically
addressed emotion regulation development in adolescence. Several recent
publications have undertaken the monumental task of comprehensively reviewing
theory relevant to emotional development
[@bussTheoriesEmotionalDevelopment2019;
@riedigerEmotionRegulationAdolescence2014]. Thanks, in part, to these efforts,
we can provide a summary review of the theoretical landscape as context for our
empirical contribution.<!--Several recent publications have undertaken the
monumental task of providing a comprehensive overview of theories of emotional
development [e.g., @bussTheoriesEmotionalDevelopment2019;
@holodynskiDevelopmentEmotionsEmotion2006;
@riedigerEmotionRegulationAdolescence2014], and of the empirical work
pertaining to emotions in adolescence
[@coe-odessEmergentEmotionsAdolescence2019].  -->

Two theories widely invoked to frame research on development, including
emotional development, are Bronfenbrenner's bioecological model, and Sameroff's
transactional model. Bronfenbrenner's model
[@bronfenbrennerBioecologicalModelHuman2007] describes how the environment
shapes individual development. Each individual is imbued with certain
biological predispositions, and develops over time, in interaction with
contextual influences. These influences range from the microsystem, composed of
those people close to the individual, to the macrosystem, consisting of
indirect political and economic influences, to the exosystem, consisting of
cultural norms and values. Sameroff's model is somewhat compatible with the
bioecological model [@sameroffUnifiedTheoryDevelopment2010], but focuses more
on the interplay between nature and nurture. It conceptualizes development as a
product of reciprocal influences between the child and the environment.
Sameroff distinguishes between proximal influences, for instance, through
parents (the microsystem in Bronfenbrenner's model), and distal influences;
structural factors indirectly shaping development, like socio-economic status,
schools, and the community (macro- and exosystem). With increasing age, distal
influences gradually gain ground on proximal influences.

Turning now towards theories specifically addressing emotion regulation in
adolescence, Hall's notion of "storm and stress" is one of the oldest [see
@arnettAdolescentStormStress1999]. It describes how hormonal changes diminish
self-control and increase reactivity, leading to difficulties in emotion
regulation, conflict with parents, and risky behavior. Recent work provides a
more nuanced perspective on this theory. For example, although the notion of
diminished self-control and increased emotional reactivity has been upheld, it
has been recast as a necessary change that facilitates emotional maturation
[@croneUnderstandingAdolescencePeriod2012], at the risk of emotional
disturbance [@arnettAdolescentStormStress1999; @leeAdolescentMentalHealth2014].
More recently, Sroufe [-@sroufeEmotionalDevelopmentOrganization1995] presented
a theory focused on normative emotional development in early childhood. It
argues that, as children grow older, their increasing self-regulatory abilities
drive a transition from external emotion regulation by primary caregivers
towards autonomous emotion regulation. This theory focuses on two drivers of
development: social and cognitive influences. Social influences mainly occur
through parental co-regulation, parenting behaviors, and parent-child
attachment. Cognitive influences occur through the development of the central
nervous system (CNS), cognition, and self-regulation. Several other theories,
introduced below, have extemporized on these sources of influence.

Several other models echo this emphasis on neorocognitive development as a driver of emotional
development, including emotion regulation development.
For example, polyvagal theory closely links emotional experience
- and regulation - to autonomous nervous system functioning
[@porgesOrientingDefensiveWorld1995]. A more recent neorocognitive theory, with
substantial relevance to adolescents' emotion regulation development, is Crone
and Dahl's model of social-affective engagement and goal flexibility
[@croneUnderstandingAdolescencePeriod2012]. This theory is based on the notion
of a "maturity gap" in middle-adolescence, arising from a developmental
asymmetry between motivational and inhibitory brain circuits
[@craccoEmotionRegulationChildhood2017;
@caseyBrakingAcceleratingAdolescent2011]. To
a greater extent than related writings, however, @croneUnderstandingAdolescencePeriod2012
set out to explain adolescents' diverging destinies; the
question why some youngsters flourish while others languish. Crone and Dahl
argue that adolescents' cognitive engagement is dynamically responsive to
social and motivational goal salience. This flexibility, on the one hand,
prepares adolescents to effectively engage cognitive systems in novel
challenging situations in a way that facilitates developing mature regulatory
abilities. On the other hand, it places them at risk to act impulsively in
pursuit of peer approval. In terms of explanans,
this theory focuses primarily on cognitive factors
and the role of peers, with less attention to factors such as parenting.

@@

Whereas our primary interest is in theories of emotional development,
theories of the phenomenon of
emotion regulation are also somewhat relevant. One influential theory is Gross'
[-@grossHandbookEmotionRegulation2013] process model, which describes the
process of emotion regulation, from eliciting cue to ultimate response.
Individuals modulate the different stages of this process using strategies,
consciously or otherwise. The effectiveness, desireability, and consequences of
different strategies depend on the cultural context [see
@bariolaChildAdolescentEmotion2011]. Similar to Gross' theory, the social
information processing theory of emotion also describe the role of intrapsychic
processes and strategies in emotion experience and reulation
[@lemeriseIntegratedModelEmotion2000].

@@

## Shortcomings of existing theory

Despite the abundance of relevant theory,
the existing literature has several limitations when the goals is to
identify which
risk factors render some adolescents more susceptible
to maladaptive emotional development than others.
One crucial shortcoming
is the lack of explicit attention to the life phase of adolescence. This is
unfortunate, as empirical research consistently indicates that adolescence is a
crucial life phase for emotion regulation development. Adolescence differs
qualitatively from infancy, childhood, and adulthood. It is therefore
questionable whether theories focused on different ages can be generalized to adolescents.
By contrast, some theories focus on adolescence specifically [e.g.,
@croneUnderstandingAdolescencePeriod2012], but only address emotion regulation
in passing.
A second, related, shortcoming is that some relevant theories lack
a developmental component <!--, and thus lack a well-substantiated
understanding of developmental processes--> [see
@bussTheoriesEmotionalDevelopment2019;
@croneUnderstandingAdolescencePeriod2012]. 
A third shortcoming is that theories
vary widely in scope: Some are very broad and all-encompassing, but somewhat
non-specific, such as the biopsychosocial models of Bronfenbrenner and
Sameroff. Others describe a specific phenomenon in great detail, but are not
embedded in a larger unifying framework. Examples include Morris' tripartite
model, Crone and Dahl's model, and polyvagal theory. It would be beneficial to
bridge these levels of analysis.
One limitation echoed across many
publications in this field is that more theory formation is required to provide
a unified framework that could guide future empirical work.
The present study
seeks to lay the groundwork for such theory development,
by providing an inductive conceptual
review of the field.

@@

Whereas theory provides a top-down understanding of an area of research,
literature reviews can provide a bottom-up, inductive understanding. Given the
noted absence of a single overarching theory, reviews might provide additional
insight into factors considered relevant in the etiology of adolescents'
emotion regulation competencies. For example, a review by
@bariolaChildAdolescentEmotion2011 addresses individual psychological factors
such as temperament, and biological factors like neurocognitive development and
genetic predispositions. Furthermore, the authors address proximal social influences,
including parents, teachers, and peers, and specific mechanisms through which
these individuals exert their influence, including parenting and modeling.
Finally, they address distal factors, including the role of culture and the
media. Barriola and colleagues also offer recommendations for future research,
<!--most notably, they call for a robust theoretical framework to guide future
research.--> calling for research on the role of parents beyond early
childhood, and including the role of fathers - a limitation which has since
been addressed by a new wave of empirical research on fathers' role in
emotional development [@vanlissaRoleFathersMothers2019;
@vanlissaMothersFathersQuantitative2020].

@@

There are notable parallels between the relevant factors highlighted by these
narrative reviews, and the aforementioned theoretical literature. Nevertheless,
these literature reviews also touch upon issues that have received little
attention in theoretical work. This highlights the value of both top-down and
bottom-up approaches to understanding the phenomenon of adolescents' emotion
regulation. One remaining limitation is that, to date, all reviews in this
field have been unstructured narrative reviews. The present publication seeks
to complement preceding work by conducting a comprehensive empirical
literature review, and mapping the factors associated with adolescent emotion
regulation.

@@

The present paper aims to develop a nomological network of the constructs
related to adolescent emotion dysregulation, based on a systematic review of
the literature. A nomological network is a diagrammatic representation of a
phenomenon; a proto-theoretical device that describes relationships between
constructs relevant to the theory [see @alaviAligningTheoryMethodology2018]. It is a precursor to a more
explicit theory of the phenomenon. The present paper uses an inductive approach
to develop this nomological network. First, we conduct a systematic search to
elicit a corpus of relevant literature. Second, we use text mining techniques
to extract relevant constructs from the corpus. Third, we use a dictionary to
pare down the extracted constructs to meaningful superordinate categories.
Fourth, we map interrelations between these constructs. Fifth, we classify
contstructs as potential predictors or outcomes.
<!--, and we categorize predictors
by level of analysis-->
<!-- (CJ: Sameroff, which level of biopsychosocial)-->.
<!-- t, using text mining to map potential predictors of adolescent emotion
dysregulation.
identify relevant constructs of adolescents’ emotion regulation development,
and moderators of parents’ role therein. The resulting insights are
consolidated into theory.
The present study intends to map the extant literature
First, I will systematically review predictors currently considered relevant in
the adolescent emotion regulation literature.  -->

@@

if(!file.exists("recs_final.csv")){
  recs_final <- unique_recs[screen$key[!screen$drop_these], ]
  write.csv(recs_final, "recs_final.csv", row.names = FALSE)
} else {
  recs_final <- read.csv("recs_final.csv")
}
```
Papers were screened based on two main criteria:
First, papers had to address emotion regulation, emotional problems, or similar.
Second, the target population must overlap with the life phase of adolescence, defined as the age range from 10-25.
Thus, for example, an article reviewing the role of hormonal changes in emotional adjustment throughout the entire life course, including adolescence, would be deemed relevant.
Screening was initially conducted in Rayyan QRCI [@Ouzzani2016].
Papers were sorted using Rayyan's ranking algorithm.
To train the algorithm on exclusion criteria, screening focused on abstracts most likely to be excluded.
After coding `r sum(screen$rayyan == "True")+sum(screen$rayyan == "False")` papers, screening continued in ASReview - a free, open source alternative to Rayyan QRCI [@van_de_schoot_rens_2020_3828293].
This program uses text mining and active learning to build a customizable machine learning model for article inclusion.
We used a naive Bayes model, and focused on exclusion of irrelevant papers.
In ASReview, an additional `r sum(table(screen$included, screen$rayyan)[, 1])` papers were screened.
Screening again focused on articles most likely to be excluded, and continued until, among the most recently screened 100 papers, only 6 were excluded.
The screening procedure is detailed in Figure \@ref(fig:prismachart).
In total, `r nrow(recs_final)` papers were retained for analysis.

@@

<!--
## Corpus description

{r, eval = FALSE}
# Plot papers per year
df_plot <- as.data.frame(table(recs_final$PY), stringsAsFactors = FALSE)
df_plot$Year = as.numeric(df_plot$Var1)
ggplot(df_plot, aes(x= Year, y= Freq))+geom_point() + geom_path(group =1)+
  theme_bw() + scale_y_log10()
-->

# Analysis 1: Author keyword mapping

```{r, warning=FALSE, message=FALSE}
source("study1.R")
if(run_everything | !file.exists("Study1_lda_dims.txt")) source("LDA_analysis.R")
lda_dims <- read_yaml("Study1_lda_dims.txt")
```
The corpus for this first analysis consisted of the author-provided keywords for the selected articles.
We extracted keywords by document, and used a dictionary
created specifically for this study to classify related terms.
The dictionary describes `r length(dict)` terms using `r length(unlist(dict))`
regular expression queries.
The lead author composed a first draft of the dictionary.
This draft was subsequently updated in two ways:
First, while writing the introduction for this paper,
key terms from the literature were added to the dictionary.
Second, by including frequently occurring uncategorized words during preliminary analyses of the data until no more frequently occurring uncategorized words were found.
An exclusion list of methodological terms and other non-substantive words was generated using the same procedure.
After applying the dictionary and exclusion filter,
and removing records with no (remaining) keywords,
the resulting corpus consisted of `r number_docs_words[1]` documents with `r number_docs_words[2]` unique words.
<!--As term-document frequency is irrelevant in the context of keywords (i.e., keywords should not occur twice within the same document, except when two synonyms have been recoded using the dictionary),
the term-document matrix was dichotomized.-->
